Should geophysicists use gravity anomaly or gravity disturbance?

### 2018-05-06 manuscript/letter.tex
review the summary and other small stuff

\begin{summary}
 The gravity anomaly has long been used by geophysicists for 
 representing the gravitational effect produced by gravity
 sources.This quantity, however, can be used in this way as a good
 approximation. Rigorously, the gravitational effect produced by the sources
 is represented by the gravity disturbance, which is a well-known 
 quantity in geodesy.
 In this paper, we discuss the fundamental concepts supporting 
 the fact that gravity disturbance is the more appropriated quantity
 to be used for estimating density variations related to geological
 bodies in subsurface. We show that the maximum absolute difference 
 between the gravity disturbance and the Free-air anomaly in Brazil 
 reaches a maximum absolute value of $\approx 9.3$ mGal. We also stress 
 that almost all authors in the literature have implicitly used the 
 gravity anomaly as an approximation of the gravity disturbance, 
 which can be represented by a harmonic function and, consequently, 
 satisfies the premise behind most of 
 the interpretation techniques used in geophysics (e.g., 
 upward/downward continuation, data processing with equivalent layer,
 conversions between gravity and magnetic data, computation of vertical 
 derivatives via Fourier and Hilbert transforms).
 Although the gravity anomaly can be considered a good approximation of the 
 gravity disturbance for most practical applications, 
 it is important to bear in mind not only the terminology 
 changes, but also the conceptual assumptions.

@@

The resultant of the virtual 
gravitational and centrifugal forces exerted by the normal
Earth on a body at rest at a point $P$ is called 
\textit{normal gravity vector} and its intensity is called 
\textit{normal gravity}.
In geodesy, any model used to represent the normal gravity field
can be arbitrarily defined for the only purpose
of keeping the difference from the actual gravity field as small 
as possible \citep{vanicek1987}.

It is worth noting that, although the normal Earth has the
same total mass (including the atmosphere) of the Earth,
its internal density distribution is unknown.
The search for physically meaningful mass distributions 
that generate a required normal gravity field
has geophysical rather than geodetic motives \citep{marussi1974}.
The only condition imposed on its internal density
distribution is that it produces a gravity field
having a particular equipotential which coincides
with its limiting surface.
For convenience, we denote any density distribution 
satisfying this condition as a \textit{normal density distribution}.

@@

In the TCS (Fig. \ref{fig:TCS}a),
the gravity disturbance $\delta g_{i}$ (eq. \ref{eq:gravity-disturbance-approx-TCS})
can be defined as the $z$-component of the gravitational attraction exerted by
the gravity sources at the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance produced by
a gravity source can be represented by the following harmonic function:
\begin{equation}
d_{i} = \partial_{z} \phi_{i} \: ,
\label{eq:gz-local}
\end{equation}
where $\partial_{z} \phi_{i}$ is the first 
derivative, evaluated at the point $(x_{i},y_{i},z_{i})$, of the function 
\begin{equation}
\phi(x,y,z) = \int\int\limits_{v}\int 
\frac{c_{g} \, G \, \Delta\rho(x^{\prime}, y^{\prime}, z^{\prime}) \: dv^{\prime}}
{\sqrt{(x - x^{\prime})^{2} + 
(y - y^{\prime})^{2} + (z - z^{\prime})^{2}}} \: ,
\label{eq:phi}
\end{equation}
with respect to the variable $z$. Is this integral, 
$G$ is the Newtonian constant of gravitation (in $m^{3} \, kg^{-1} \, s^{-2}$),
$c_{g} = 10^{5}$ is a constant factor transforming from $m \, s^{-2}$ 
to milligal (mGal), $\Delta\rho(x^{\prime}, y^{\prime}, z^{\prime})$
is the density contrast (in $kg \, m^{-3}$) at a point 
$(x^{\prime}, y^{\prime}, z^{\prime})$ within 
the volume $v$ of the gravity source and the integration is conducted 
over $x^{\prime}$, $y^{\prime}$ and $z^{\prime}$.
This equation can be easily generalized for the case
of multiple gravity sources.
Practically, all articles accomplishing gravity modelling
use the quantity $d_{i}$ (eq. \ref{eq:gz-local})
to represent the gravity anomaly produced by 
gravity sources \citep[e.g.,][]{blakely1996}. 
Consequently, almost all geophysicists use the gravity
anomaly as an approximation of the gravity disturbance
produced by the gravity sources. Notice that $d_{i}$ 
(eq. \ref{eq:gz-local}) does not depend on the
height with respect to the Geoid (orthometric height). 
Rather, it depends on the relative position 
of the observation points $(x_{i}, y_{i}, z_{i})$ with respect to
the gravity sources in the TCS (Fig. \ref{fig:TCS}a).
Finally, it is important to notice that, contrary to what 
is written in some basic texts, gravimeters do not measure 
the quantity $d_{i}$ (eq. \ref{eq:gz-local}).

>>>>

\begin{summary}
 In applied geophysics, the gravity anomaly has long been used for representing the
 gravitational effect produced by gravity sources, which are defined as the density
 contrasts between the internal density distribution of the Earth and the normal Earth.
 The gravity anomaly is defined by the difference between the gravity (produced by the
 Earth) on the Geoid and the normal gravity (produced by the normal Earth)
 on reference ellipsoid, both at the same geodetic latitude and longitude.
 Because these quantities are not defined at the same point, 
 the gravity anomaly does not represent the gravitational effect due to the gravity
 sources and, consequently, cannot be considered a harmonic function.
 On the contrary, the gravity disturbance is defined by the difference between
 the gravity and normal gravity at the same point. Consequently, the centrifugal 
 effects can be neglected, the gravity disturbance can be considered a harmonic 
 function representing the gravitational effect due to the gravity sources and then 
 it satisfies the premise behind most of the processing techniques of potential-field
 data (e.g., upward/downward continuations). Then, rigorously, the gravity disturbance
 is  more appropriated to be used for estimating density variations in subsurface.
 The use the gravity anomaly for most of the geophysicists carries the
 implicit meaning that it can be considered a good approximation of the gravity
 disturbance. It is important to bear in mind, however, that in Brazil, for example, 
 the maximum absolute difference between the gravity disturbance (on or close 
 to the Earth’s surface) and the Free-air anomaly reaches 9.3 mGal.

@@

The resultant of the virtual 
gravitational and centrifugal forces exerted by the normal
Earth on a body at rest at a point $P$ is called 
\textit{normal gravity vector} and its intensity is called 
\textit{normal gravity}.
In geodesy, any model used to represent the normal gravity field
can be arbitrarily defined for the only purpose
of keeping the difference from the actual gravity field as small 
as possible \citep{vanicek1987}.
It is worth noting that, although the normal Earth has the
same total mass (including the atmosphere) of the Earth,
its internal density distribution is unknown.
The search for physically meaningful mass distributions 
that generate a required normal gravity field
has geophysical rather than geodetic motives \citep{marussi1974}.
The only condition imposed on its internal density
distribution is that it produces a gravity field
having a particular equipotential which coincides
with its limiting surface.
For convenience, we denote any density distribution 
satisfying this condition as a \textit{normal density distribution}.

@@

In the TCS (Fig. \ref{fig:TCS}a),
the gravity disturbance $\delta g_{i}$ (eq. \ref{eq:gravity-disturbance-approx-TCS})
can be defined as the $z$-component of the gravitational attraction exerted by
the gravity sources at the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance produced by
a gravity source can be represented by the following harmonic function:
\begin{equation}
d_{i} = \int\int\limits_{v}\int 
\frac{c_{g} \, G \, \Delta\rho(x^{\prime}, y^{\prime}, z^{\prime}) 
\, (z_{i} - z^{\prime}) \: dv^{\prime}}
{\sqrt{(x_{i} - x^{\prime})^{2} + 
(y_{i} - y^{\prime})^{2} + (z_{i} - z^{\prime})^{2}}} \: ,
\label{eq:gz-local}
\end{equation}
where $G$ is the Newtonian constant of gravitation (in $m^{3} \, kg^{-1} \, s^{-2}$),
$c_{g} = 10^{5}$ is a constant factor transforming from $m \, s^{-2}$ 
to milligal (mGal), $\Delta\rho(x^{\prime}, y^{\prime}, z^{\prime})$
is the density contrast (in $kg \, m^{-3}$) at a point 
$(x^{\prime}, y^{\prime}, z^{\prime})$ within 
the volume $v$ of the gravity source and the integration is conducted 
over $x^{\prime}$, $y^{\prime}$ and $z^{\prime}$.
This equation can be easily generalized for the case
of multiple gravity sources.
Practically, all articles accomplishing gravity modelling
use the quantity $d_{i}$ (eq. \ref{eq:gz-local})
to represent the gravity anomaly produced by 
gravity sources \citep[e.g.,][]{blakely1996}. 
Consequently, almost all geophysicists use the gravity
anomaly as an approximation of the gravity disturbance
produced by the gravity sources. Notice that $d_{i}$ 
(eq. \ref{eq:gz-local}) does not depend on the
height with respect to the Geoid (orthometric height). 
Rather, it depends on the relative position 
of the observation points $(x_{i}, y_{i}, z_{i})$ with respect to
the gravity sources in the TCS (Fig. \ref{fig:TCS}a).
Finally, it is important to notice that, contrary to what 
is written in some basic texts, gravimeters do not measure 
the quantity $d_{i}$ (eq. \ref{eq:gz-local}).
